---
title: "Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(BOSSS)
```


## Neil's example

```{r}
source("C:/Users/meddwilb/OneDrive - University of Leeds/Documents/Research/Projects/BOSSS example/PowCalc_Binary_DGM_5_freq.R")
source("C:/Users/meddwilb/OneDrive - University of Leeds/Documents/Research/Projects/BOSSS example/PowCalc_Binary_Analysis_5_freq.R")
source("C:/Users/meddwilb/OneDrive - University of Leeds/Documents/Research/Projects/BOSSS example/PowCalc_Binary_DGM_5a_freq.R")
source("C:/Users/meddwilb/OneDrive - University of Leeds/Documents/Research/Projects/BOSSS example/PowCalc_Binary_BPPcalculator_5_freq.R")
source("C:/Users/meddwilb/OneDrive - University of Leeds/Documents/Research/Projects/BOSSS example/PowCalc_Binary_Design_5_freq.R")

# n = Total sample size
# n_interim = Interim sample size
# p0_true and p1_true:= the assumed true probabilities of an event in the control (0) and experimental (1) arms
# alpha = the threshold used at the final analysis to determine whether the result is significant or not. At final analysis, the bounds of the 100*(1 - alpha/2)% credible interval of the log-odds ratio is compared to 0.
# BPP_numsims:= how many simulations are used to evaluate Bayesian Predictive Power at the interim analysis
# BPP_threshold:= BPP must be >= this value in order for the trial to continue to final analysis
# Remaining: the analysis of the trial assumes beta prior distributions on p0 and p1. p0 ~ beta(APr_p0_alpha, APr_p0_beta) and p1 ~ beta(APr_p1_alpha, APr_p1_beta)
# by default these are both set to be beta(1,1), non-informative.

sim_trial <- function(design = c(400, 200, 0.2, 0.05), hypothesis = c(0.2, 0.1)) {
  
  design <- as.numeric(design); hypothesis <- as.numeric(hypothesis)
  
  n <- design[1]; n_interim <- design[2]; BPP_threshold <- design[3]; alpha <- design[4]
  p0_true <- hypothesis[1]; p1_true <- hypothesis[2]
  
  BPP_numsims=200; APr_p0_alpha=1; APr_p0_beta=1; APr_p1_alpha=1; APr_p1_beta=1
  
  # Generate a dataset
  chk_DGM_output <- DGM_run(n, n_interim, p0_true, p1_true)
  #Debugging
  # print("DGM output DONE")
  
  # Analyse
  chk_Analysis_output <- Analysis_run(compiled_model, interim_data=chk_DGM_output$interim_data, chk_DGM_output$full_data, alpha, BPP_numsims, APr_p0_alpha, APr_p0_beta, APr_p1_alpha, APr_p1_beta)
  #Debugging
  # print("Analysis output DONE")
  BPP_sig <- (chk_Analysis_output$BPP >= BPP_threshold)
  final_sig <- (chk_Analysis_output$beta1quants[2] <= 0 | chk_Analysis_output$beta1quants[1] >= 0)
  trial_sig <- (BPP_sig & final_sig)
  trial_non_sig <- !trial_sig
  N_trial <- n_interim + BPP_sig*(n-n_interim)
  
  # Return the results
  results <- c(
    # BPP_sig = BPP_sig,
    trial_sig = trial_sig,
    trial_non_sig = trial_non_sig,
    N_trial = N_trial
  )
  
  # Return the results
  # results <- c(
  #   beta1_est_int = chk_Analysis_output$beta1mean_interim,
  #   beta1_lower_int = chk_Analysis_output$beta1quants_interim[1],
  #   beta1_upper_int = chk_Analysis_output$beta1quants_interim[2],
  #   beta0_est_int = chk_Analysis_output$beta0mean_interim,
  #   BPP = chk_Analysis_output$BPP,
  #   BPP_sig = BPP_sig,
  #   beta1_est_final = chk_Analysis_output$beta1mean,
  #   beta1_lower_final = chk_Analysis_output$beta1quants[1],
  #   beta1_upper_final = chk_Analysis_output$beta1quants[2],
  #   beta0_est_final = chk_Analysis_output$beta0mean,
  #   final_sig = final_sig,
  #   trial_sig = trial_sig,
  #   trial_non_sig = trial_non_sig,
  #   N_trial = N_trial
  # )
  
  return(results)
}  

# For example,
sim_trial()
```


```{r}
design_space <- design_space(name = c("n", "n_interim", "BPP_threshold", "alpha"),
                             lower = c(700,300,0,0), 
                             upper = c(1400,650,1,0.2))
```


```{r}
hypotheses <- hypotheses(par_name = c("p0_true", "p1_true"),
                         values = matrix(c(0.12, 0.12,
                                           0.12, 0.06), ncol = 2),
                         hyp_names = c("null", "alt"))
```


```{r}
constraints <- constraints(name = c("a", "b"),
                   out = c("trial_sig", "trial_non_sig"),
                   hyp = c("null", "alt"),
                   nom = c(0.2, 0.4),
                   delta =c(0.95, 0.95),
                   stoch = c(TRUE, TRUE))
```


```{r}
objectives <- objectives(name = c("f1", "f2", "f3"),
                 out = c("trial_sig", "trial_non_sig", "N_trial"),
                 hyp = c("null", "alt", "null"),
                 weight = c(100, 100, 1),
                 stoch = c(TRUE, TRUE, TRUE),
                 binary = c(TRUE, TRUE, FALSE))
```

We now put this simulation function and set of data frames together to create an object of class `BOSSS_problem`. 

```{r}
prob <- BOSSS_problem(sim_trial, design_space, hypotheses, constraints, objectives)
```

### Initialisation


```{r}
size <- 40
N <- 50

sol <- BOSSS_solution(size, N, prob)

print(sol)
plot(sol)
```


### Iteration

```{r}
for(i in 1:10) {
  sol <- iterate(sol, prob, N) 
}
print(sol)
plot(sol)
```


### Diagnostics

```{r}
# Pick a specific design from the Pareto set
design <- sol$p_set[1,]

plots <- one_d_plots(design, prob, sol)

for(i in 1:length(plots)) {
  print(plots[[i]])
}
```

```{r}
r <- check_point(design, prob, sol, N=50) 
```

```{r}
r <- check_point(design, prob, sol, N=50, r) 
```

## Frequentist-Bayesian 

Suppose we plan a Bayesian analysis of a trial comparing two therapies, where we expect a multiple membership structure. In particular, each of the $2n$ patients in the trial will receive six treatment sessions, but these sessions can be delivered by any of the $m$ therapists in their arm of the trial. Considering the number of therapists in each arm as fixed at $m = 10$, we want to know how many patients are required for the trial to have sufficient power whilst controlling the type I error rate (both being frequentist concepts).

We start with the simulation. To simulate the data we first construct the matrix $\mathbf{V}$ (`v_m` in the code) which has a row for every patient and a colum for every therapist, with $\mathbf{V}_{ij}$ denoting the proportion of patient $i$'s treatment delivered by therapist $j$, with $i = 1, \ldots , 2n$ and $j = 1, \ldots , 2m$. Denoting within-cluster variance by $\sigma_e^2$ (`var_e`), between-cluster variance by $\sigma_v^2$ (`var_v`), and the ICC by $\rho$ (`rho`), the outcome data can be simulated as

$$
\mathbf{y} \sim N\left(E[\mathbf{y}], \sigma_e^2 \mathbf{I}_{2n} + \sigma_v^2 \mathbf{V} \mathbf{V}^T  \right).
$$
This data is then analysed using Stan via `brm

```{r}
library(dirmult)

sim_trial <- function(n=100, var_e=0.9, rho=0.1, mu=0.3){
  
  v_0 <- cbind(t(rmultinom(n = n, size = 6, prob = rdirichlet(1, rep(0.7,m)))),
               matrix(rep(0, n*m), nrow = n))
  v_1 <- cbind(matrix(rep(0, n*m), nrow = n),
               t(rmultinom(n = n, size = 6, prob = rdirichlet(1, rep(0.7,m)))))
  
  v_m <- rbind(v_0, v_1)
  
  var_v <- rho*var_e/(1-rho);
  
  Sigma <-  var_e * diag(2*n) #+ var_v * (v_m %*% t(v_m)) 
  
  X <- matrix(c(rep(1, 2*n), rep(0:1, each = n)), ncol = 2)
  
  y <- t(mvtnorm::rmvnorm(1, mean = X %*% c(0, mu), sigma = Sigma))
  
  df <- data.frame(y = y,
                   trt = X[,2])
  
  v_m <- as.data.frame(v_m)
  
  df <- cbind(df, t(apply(v_m, 1, function(x) rep(names(x), x))))
  names(df)[3:8] <- paste0("v", 1:6)
  
  fit_empty <- brm(y ~ trt, #+ (1| mm(v1, v2, v3, v4, v5, v6)),
           prior = prior(normal(0, 5), class = b),
           data = df,
           chains = 0,
           silent = 2,
           file = "fit_empty")
  
  fit_empty <- readRDS("fit_empty.rds")
  
  fit <- update(fit_empty, 
                newdata = df,
                chains = 1)
  
  g <- fixef(fit)[2,3] > 0
  s <- !g
  
  return(c(g = g, s = s))
}

sim_trial()
```
