---
title: "Notes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Notes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(BOSSS)
library(ggplot2)
```

A place to build up notes and examples for a future tutorial-style document.

Note - a Twitter thread on power analysis had a couple of people mentioning the R package simr for power through simulation when using mixed models. So might be worth including that in one of our examples, to save people having to write their own simulation code?

# Examples

## Adaptive design (NIFTy)

### Problem

Simulation:
```{r}
#n is the total sample size
#ninterim is the number of patients at the interim analysis (proportion)
#ainterim is alpha at interim analysis (threshold p-value at interim analysis)
#afinal is alpha at final analaysis (threshold p-value for 2nd and final analysis)
#this means overall alpha for the trial is ainterim*afinal
#
#pcontshort is the probability of 1 day PoSH in the control arm
#pexpshort is the probability of 1 day PoSH in the experimental arm
#pcontlong is the probability of 6 month PoSH in the control arm
#pexplong is the probability of 6 month PoSH in the experimental arm

sim_trial <- function(design = c(300, 0.5, 0.4, 0.1),
                      hypothesis = c(0.25, 0.125, 0.1, 0.03))
{
  design <- as.numeric(design); hypothesis <- as.numeric(hypothesis)
  
  n <- design[1]; ninterim <- design[2]
  ainterim <- design[3]; afinal <- design[4]
  
  pcontshort <- hypothesis[1]; pexpshort <- hypothesis[2]
  pcontlong <- hypothesis[3]; pexplong <- hypothesis[4]
  
  ninterim <- floor(ninterim*n)
  
  patients<-c(1:n) #create patients
  treat<-rep(c(1,2),ceiling(n/2))[1:n] #assign to treatment 1 or 2
  
  short<-rep(0,n) #short term outcome
  long<-rep(0,n) #long term outcome
  
  data<-data.frame(patients,treat,short,long) #combine into dataset
  
  #generate result 1/0 for short term outcome. If short term outcome=0 then long term outcome=0
  #If short term outcome=1 then long term outcome has probability pcontlong/pcontshort
  #repeat for treatment=2
  for(i in 1:n){
    if(treat[i]==1){
      data$short[i]<-rbinom(1,1,pcontshort)
      if(data$short[i]==0){
        data$long[i]<-0
      }
      if(data$short[i]==1){
        data$long[i]<-rbinom(1,1,pcontlong/pcontshort)
      }
    }
    else{
      data$short[i]<-rbinom(1,1,pexpshort)
      if(data$short[i]==0){
        data$long[i]<-0
      }
      if(data$short[i]==1){
        data$long[i]<-rbinom(1,1,pexplong/pexpshort)
      }
    }
  }
  #perform chi squared test on short term outcome for 1st ninterim patients
  data2<-data[1:ninterim,]
  tbl<-table(data2$short,data2$treat)
  test<- suppressWarnings(chisq.test(tbl)$p.value) #get p-value
  
  #if p<ainterim, perform chi squared test on long term outcome for all patients
  if(test<ainterim){
    tbl2<-table(data$long,data$treat)
    test2<- suppressWarnings(chisq.test(tbl2)$p.value)
  }
  
  #if p>ainterim, trial unsuccessful at interim and final analysis
  #if p<ainterim, trial successful at interim:
  #if p2>afinal, trial unsuccessful at final analysis
  #if p2<afinal, trial successful at final analysis
  if(test>=ainterim){
    return(c(g = 0, s = 1, n = ninterim))
  }
  if(test<ainterim){
    if(mean(data2[data2$treat == 1,"short"]) < mean(data2[data2$treat == 2,"short"])){
      return(c(g = 0, s = 1, n = ninterim))
    } else {
      if(test2>afinal){
        return(c(g = 0, s = 1, n = n))
      }
      if(test2<afinal){
        return(c(g = 1, s = 0, n = n))
      }
    }
  }
}

# For example,
sim_trial()
```

Design space and hypotheses:
```{r}
design_space <- design_space(name = c("n", "ninterim", "ainterim", "afinal"),
                             lower = c(100,0.05,0,0), 
                             upper = c(500,0.5,1,1))

hypotheses <- hypotheses(name = c("pcontshort", "pexpshort", "pcontlong", "pexplong"),
                         param_matrix = matrix(c(0.25, 0.25, 0.25, 0.25, 
                                                 0.25, 0.125, 0.1, 0.03), ncol = 2))
```


### Initialisation

Constraints and objectives:

```{r}
constraints <- constraints(name = c("a", "b"),
                   out_i = c(1, 2),
                   hyp_i = c(1, 2),
                   nom = c(0.2, 0.2),
                   delta =c(0.95, 0.95),
                   stoch = c(TRUE, TRUE))

objectives <- objectives(name = c("f1", "f2", "f3"),
                 out_i = c(1, 2, 3),
                 hyp_i = c(1, 2, 1),
                 weight = c(100, 100, 1),
                 stoch = c(TRUE, TRUE, TRUE))

prob <- BOSSS_problem(sim_trial, design_space, hypotheses, constraints, objectives)
```

Initialise solution:

```{r}
size <- 40
N <- 100

sol <- BOSSS_solution(size, N, prob)

print(sol)
plot(sol)
```


### Iteration

```{r}
for(i in 1:5) {
  sol <- iterate(sol, prob, N) 
}
print(sol)
plot(sol)
```

### Diagnostics

To check the GP models are giving sensible predictions, we can choose a specific design and then plot the predictions for each model along the range of each design variable.

```{r}
# Pick a specific design from the Pareto set
design <- sol$p_set[1,]

plots <- one_d_plots(design, prob, sol)

for(i in 1:length(plots)) {
  print(plots[[i]])
}
```

When we've chosen our design, we might want to check its properties by running computing more MC estimates, which we can also compare against the model predictions. 

```{r}
r <- check_point(design, prob, sol, N=10^2) 
```

If we decide we want to run more simulations at the same point, we can pass the previous results in so they are built upon.

```{r}
r <- check_point(design, prob, sol, N=10^2, r) 
```

### Presentation

Plotly is useful to look at the specific points on the Pareto front to choose amongst them. To do - add an identifier so any point on the Pareto front can be easily linked to the corresponding design in the Pareto set.

```{r}
# Plot the Pareto front as a plotly object if installed
is_plotly_available <- require(plotly)
if(is_plotly_available) {
  ggplotly(plot(sol))
}
```

### Sensitivity analysis

Use a generalised hypothesis defined by ranges (which may be degenerate point values) for each parameter. Take a suggested design and fit new models to cover these ranges.

## Simple two-sample t-test

To do.

```{r}
sim_trial <- function(design, hypothesis){
  # Exctract the design variables and parameters
  design <- as.numeric(design); hypothesis <- as.numeric(hypothesis)
  n <- design[1]; k <- design[2]
  mu <- hypothesis[1]; var_u <- hypothesis[2]; var_e <- hypothesis[3]
  
  # Number of patients in each cluster
  m <- n/k
  # SD of cluster means
  s_c <- sqrt(var_u + var_e/m)
  # Simulate cluster means
  x0 <- stats::rnorm(k, 0, s_c); x1 <- stats::rnorm(k, mu, s_c)
  
  return(c(s = stats::t.test(x0, x1)$p.value >= 0.05, n = n, k = k))
}

det_obj <- function(design){
    #o <- matrix(design, ncol = 2)[,1:2]
    #c(s = NA, p = o[1], c = o[2])
  cbind(NA, design)
}

design_space <- data.frame(name = c("n", "k"),
                 low = c(50, 5),
                 up = c(500, 50),
                 int = c(TRUE, TRUE))

hypotheses <- data.frame(mu = c(0.3, 0),
                   var_u = c(0.05, 0.05),
                   var_e = c(0.95, 0.95))

constraints <- data.frame(name = "a",
                   out_i = 1,
                   hyp_i = 2,
                   nom = 0.2,
                   delta = 0.975,
                   stoch = TRUE)

objectives <- data.frame(name = c("f1", "f2", "f3"),
                 out_i = c(2, 3, 1),
                 hyp_i = c(1, 1, 1),
                 weight = c(2, 5, 50),
                 stoch = c(FALSE, FALSE, TRUE))
```

## Bayesian / decision theoretic multilevel model

Note that the method extends trivially to any kind of Bayesian analysis. We just use the hypotheses parameters to describe hyperparameters for the priors we want to simulate under, and use a Bayesian analysis (or freq if using a hybrid approach). This would include a decision-theoretic approach, where the analysis would give us a utility and we would use the expectation of this as an objective to maximise. One difficulty would be avoiding model compilation when using Stan or brms. We could potentially deal with this by storing a pre-compiled model as a file, and then loading and updating it for each iteration.

## Multi-state models



